### 2017-08-05 Source/Examples/form_triplet_motifs.Rmd
Add all current files

---
title: "Form Triplet Motifs"
author: "Ignatius Pang"
date: "8 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph)
library(plyr)
library(dplyr)
library(purrr)
library(tibble)
library(tidyr)

## Load the root base directory for the entire project
source( "./Common/parameters_file.R")

data_directory <- file.path( base_directory, "Data/Triplet_Motifs_R_data")

load( file.path(data_directory, "network_data_library.Rdata"))

source( file.path(base_directory, "Source/Common", "count_triplet_motifs_helper.R"))

```

```{r}

# Genetic interactions network table

filtered_costanzo_stringent <- filtered_costanzo_stringent_2016 %>% 
								dplyr::select( one_of( c( "query_oln_id_edited", "array_oln_id_edited", "genetic_interaction_score"))) %>%
								dplyr::rename( node_a = query_oln_id_edited, node_b= array_oln_id_edited )

# Re-arrange the edges such that the ID of node A is always greater than ID of node B
# This removes the symmetry in the triplet motif dataset 
a_gt_b <- dplyr::filter( filtered_costanzo_stringent, node_a >= node_b   )

b_gt_a <- dplyr::filter( filtered_costanzo_stringent, node_b > node_a   ) %>%
		  dplyr::rename( node_a=node_b, node_b=node_a )

# If the genetic interaction has been screen more than once, get the least negative score for that interaction
filtered_costanzo_stringent <- dplyr::union( a_gt_b, b_gt_a) %>%
							   dplyr::group_by( node_a, node_b) %>%
							   dplyr::summarise(max_gi_score = max(genetic_interaction_score) ) %>%
							   dplyr::rename( genetic_interaction_score=max_gi_score) %>%
							   dplyr::ungroup()

head( filtered_costanzo_stringent )

# Count the unique number of proteins in triplet motifs
length ( unique(  c(as.vector(as.data.frame(filtered_costanzo_stringent)[, "node_a"]) , 
					as.vector (as.data.frame( filtered_costanzo_stringent)[, "node_b"] )) ) )

```

```{r}
# All other biological networks, protein-protein, kinase-substrate, transcription factor-target gene
head( interactions_combined)

# Rename the tables
interactions_combined_1 <- interactions_combined %>%
							dplyr::rename( node_a = oln_id_a, 
										   node_c= oln_id_b, 
										   type_ac_long = interaction_type,
										   type_ac = interaction_type_abbrev)

interactions_combined_2 <- interactions_combined %>%
							dplyr::rename( node_b = oln_id_a, 
										   node_c= oln_id_b, 
										   type_bc_long = interaction_type,
										   type_bc = interaction_type_abbrev )

```

Total number of proteins in the integrated network
```{r}

list_of_proteins_genetic_interactions <-	filtered_costanzo_stringent  %>% 
	tidyr::gather( "node", "oln_id", 1:2) %>%
	dplyr::select ( -node) %>%
	dplyr::select ( -genetic_interaction_score) %>%
	distinct()

list_of_proteins_physical_interactions <- interactions_combined  %>% 
	tidyr::gather( "node", "oln_id", 1:2) %>%
	dplyr::select ( -node) %>%
	dplyr::select ( -interaction_type) %>%
	dplyr::select ( -interaction_type_abbrev) %>%
	distinct()

total_number_of_proteins_integrated_biological_newtork <-    list_of_proteins_genetic_interactions %>% 
	rbind ( list_of_proteins_physical_interactions ) %>%
	distinct () %>%
	count()

total_number_of_proteins_integrated_biological_newtork

```

```{r}
## Join the networks together to find triplets where at least one edge is a negative genetic interaction

				  # Join the genetic interaction to the other biological networks
triplet_motifs <- dplyr::inner_join( filtered_costanzo_stringent, interactions_combined_1, by="node_a" )  %>%
						# Closing the triplets with a third edge from the other biological networks 	
					    dplyr::inner_join( interactions_combined_2, by=c("node_b"= "node_b", "node_c"= "node_c") ) %>%
						# Select the corresponding columns
					    dplyr::select ( one_of( c("node_a", "node_b", "node_c", "type_ac", "type_bc", 
					    						  "type_ac_long", "type_bc_long", "genetic_interaction_score"))) %>%

						# Remove special cases where Node A == Node C or Node B == Node C
						dplyr::filter ( node_a != node_c) %>%
						dplyr::filter ( node_b != node_c) %>%

						# Remove duplicates
					    dplyr::distinct() %>%

						# Arrange the columns in the order that I want 
					    dplyr::arrange( node_a, node_b, node_c, type_bc, type_ac)

triplet_motifs

## Count the number of motifs
nrow( triplet_motifs)

triplet_motifs %>%
	dplyr::select( one_of ( c( "node_a", "node_b", "node_c", "type_ac", "type_bc", "genetic_interaction_score"))) %>%
	write.table ( file.path( list_of_triplets_directory,  "triplet_motifs_costanzo_2016.txt"), sep="\t", quote=FALSE, row.names = FALSE )

```

```{r}

# Count the number of each type of triplet motifs

# Count the cases where type_ac is greater than type_bc (e.g. string comparison)
a_ge_b <- filter( triplet_motifs, type_ac >= type_bc ) %>%
		group_by( type_ac, type_bc) %>%
		summarise(counts=n())

# Count the cases where type_bc is greater than type_ac (e.g. string comparison)
b_gt_a <- filter( triplet_motifs, type_bc > type_ac ) %>%
		group_by( type_ac, type_bc) %>%
		summarise(counts=n()) %>%
		rename ( type_ac= type_bc, type_bc = type_ac) %>%
		select( one_of( c("type_ac", "type_bc", "counts") ))

# Merge the above results together
triplet_motif_counts <- dplyr::union( a_ge_b, b_gt_a)  %>%
		group_by( type_ac, type_bc) %>% 
		summarise(total_count=sum(counts)) %>%
		arrange( type_ac, type_bc)

# Here is the final results
triplet_motif_counts

dim( triplet_motif_counts)

triplet_motif_counts <- as.data.frame(triplet_motif_counts)

triplet_motif_counts  <- mutate( triplet_motif_counts, motif_type = paste(type_ac, type_bc, sep="") )  %>%
						arrange ( desc(total_count))

triplet_motif_counts[, "motif_type"] <- factor ( as.vector(triplet_motif_counts[,"motif_type"]), levels=as.vector(triplet_motif_counts[,"motif_type"]))

library(ggplot2)

ggplot ( triplet_motif_counts, aes(motif_type, total_count)) + 
	geom_point()

```

Find the number of edges of each type in the triplet motifs network
```{r}

ab_interactions <- triplet_motifs %>%
					dplyr::select( one_of(c( "node_a", "node_b")))  %>%
					dplyr::rename( node_x = node_a, node_y = node_b) %>% 
					dplyr::mutate( type="gi")

ac_interactions <- triplet_motifs %>% 
	dplyr::select( one_of(c( "node_a", "node_c", "type_ac"))) %>%
	rename( type= type_ac, node_x = node_a, node_y = node_c)

bc_interactions <- triplet_motifs %>% 
	dplyr::select( one_of(c( "node_b", "node_c", "type_bc"))) %>%
	rename( type= type_bc, node_x = node_b, node_y = node_c)

# This function makes sure all of the interactions are in the same direction.
# This helps with removing duplicates. 
fix_interaction <- function ( interactions_list) {

	interactions_list <- setNames( as.list(as.vector(interactions_list) ), names( interactions_list) )

	if( ( interactions_list$type == 'p'  | interactions_list$type == 'gi'  | interactions_list$type == '') 
		& interactions_list['node_x'] < interactions_list$node_y ) {
		# perform swapping
		temp <- interactions_list$node_y
		interactions_list$node_y <- interactions_list$node_x
		interactions_list$node_x <- temp
	} else if  ( interactions_list$type == 'tu' ) {

		temp <- interactions_list$node_y
		interactions_list$node_y <- interactions_list$node_x
		interactions_list$node_x <- temp
		interactions_list$type <- "td"
	} else if  ( interactions_list$type == 'ku' ) {

		temp <- interactions_list$node_y
		interactions_list$node_y <- interactions_list$node_x
		interactions_list$node_x <- temp
		interactions_list$type <- "kd"
	}

	return( unlist( interactions_list )) ## Need to unlist it
}

ab_interactions_list <- as.list(as.data.frame(t(ab_interactions))) %>%
						map(fix_interaction)

ab_interactions <- ab_interactions_list %>%
					ldply() %>%    ## Convert back to a data.frame
					select ( -(.id)) %>% ## remove the .id column 
					distinct()

ac_interactions_list <- as.list(as.data.frame(t(ac_interactions))) %>%
						map( fix_interaction)

ac_interactions <- ac_interactions_list %>%
					ldply() %>% 
					select ( -(.id))

bc_interactions_list <- as.list(as.data.frame(t(bc_interactions))) %>%
						map( fix_interaction)

bc_interactions <- bc_interactions_list %>%
					ldply() %>% 
					select ( -(.id))

interactions_table <- bind_rows( ac_interactions, bc_interactions, ab_interactions) %>%
					  distinct()

## Total number of unique interactions in the network of triplets. Removing overlapping edges. Ignoring direction of edges
total_number_of_interactions_triplet_network_no_overlap_no_direction <- interactions_table %>% 
												mutate ( type = '') %>%
											    t() %>% as.data.frame() %>% as.list() %>%
												map( fix_interaction) %>%
												ldply() %>% 
												select ( -(.id)) %>% 
												distinct() %>%
												count()
total_number_of_interactions_triplet_network_no_overlap_no_direction

## Number of each type of interactions 
interactions_count_table <- interactions_table %>%
							group_by(type) %>%
							summarise ( num_interactions = n() )

### Total number of proteins 
total_number_of_proteins_in_triplets <-	interactions_table  %>% 
	tidyr::gather( "node", "oln_id", 1:2) %>%
	dplyr::select ( -node) %>%
	dplyr::select ( -type) %>%
	distinct() %>% 
	count()

### Number of proteins in each type of network
all_nodes <- interactions_table %>% 
	tidyr::gather( "node", "oln_id", 1:2) %>%
	dplyr::select ( -node) %>%
	distinct() %>% 
	group_by(type) %>% 
	count()

## Number of kinases in the triplet netwrok
interactions_table %>% 
	dplyr::filter ( type=='kd') %>%
	dplyr::select ( one_of ( c( 'node_x'))) %>%
	distinct()  %>%
	count()

## Number of transcription factors in the triplet netwrok
interactions_table %>% 
	dplyr::filter ( type=='td') %>%
	dplyr::select ( one_of ( c( 'node_x'))) %>%
	distinct()  %>%
	count()

```

Total number of interactions in the integrated network
```{r}

part_a <- filtered_costanzo_stringent %>%
	dplyr::select ( -genetic_interaction_score) %>%
	rename( node_x = node_a, node_y = node_b)

part_b <- interactions_combined %>%
	dplyr::select ( one_of(c('oln_id_a', 'oln_id_b'))) %>%
	rename( node_x = oln_id_a, node_y = oln_id_b)

part_a %>% rbind(part_b) %>% 
												mutate ( type = '') %>%
											    t() %>% as.data.frame() %>% as.list() %>%
												map( fix_interaction) %>%
												ldply() %>% 
												select ( -(.id)) %>% 
												distinct() %>%
												count()

total_number_of_interactions_triplet_network_no_overlap_no_direction <- interactions_table %>% 
												mutate ( type = '') %>%
											    t() %>% as.data.frame() %>% as.list() %>%
												map( fix_interaction) %>%
												ldply() %>% 
												select ( -(.id)) %>% 
												distinct() %>%
												count()
total_number_of_interactions_triplet_network_no_overlap_no_direction

```

Genetic interactions, number of edges, number of proteins
```{r}

filtered_costanzo_stringent_2016_cleaned <- filtered_costanzo_stringent_2016 %>%
	dplyr::select ( query_oln_id_edited, array_oln_id_edited) %>% 
					dplyr::rename( node_x=query_oln_id_edited, node_y = array_oln_id_edited) %>%
					dplyr::mutate( type="gi") %>%
					as.tibble() %>%
					t() %>%
					as.data.frame() %>%
					as.list() %>%
					map(fix_interaction) %>%
					ldply() %>% 
					dplyr::select ( -(.id)) %>%
					distinct()

# Get number of unique edges	
dim( filtered_costanzo_stringent_2016_cleaned)

# Get number of proteins
length( unique( c(filtered_costanzo_stringent_2016_cleaned$node_x,  filtered_costanzo_stringent_2016_cleaned$node_y) ))

```

Kinase substrate network 
```{r}

## Number of interactions in the kinase-substrate network
kinase_network_cleaned <- kinase_network_collated %>% 
	dplyr::filter( interaction_type_abbrev == 'kd') %>%
	dplyr::select( one_of(c('oln_id_a', 'oln_id_b'))) %>%
	distinct()

dim( kinase_network_cleaned)

## count the number of kinases
kinase_list <- kinase_network_cleaned %>%
	dplyr::select ( one_of ( c ( 'oln_id_a'))) %>%
	dplyr::rename( oln_id = oln_id_a) %>%
	distinct()

count(kinase_list )

## count the number of substrates

targets_list <- kinase_network_cleaned %>%
	dplyr::select ( one_of ( c ( 'oln_id_b'))) %>%
	dplyr::rename( oln_id = oln_id_b) %>%
	distinct()

substrates_excluding_kinases <- setdiff( targets_list, kinase_list)

count( substrates_excluding_kinases)

```

Protein interaction network
```{r}

sbi_interactome_cleaned <- sbi_interactome_collated %>% 
	dplyr::filter( oln_id_a > oln_id_b ) %>%
	dplyr::select( one_of(c('oln_id_a', 'oln_id_b'))) %>%
	distinct()

dim( sbi_interactome_cleaned)

length( unique( c(sbi_interactome_cleaned$oln_id_a ,  sbi_interactome_cleaned$oln_id_b )) )

```

Transcriptional regulatory network
```{r}

tf_network_collated

## Number of interactions in the transcriptional regulatory network
tf_network_cleaned <- tf_network_collated %>% 
	dplyr::filter( interaction_type_abbrev == 'td') %>%
	dplyr::select( one_of(c('oln_id_a', 'oln_id_b'))) %>%
	distinct()

dim( tf_network_cleaned)

## count the number of kinases
tf_list <- tf_network_cleaned %>%
	dplyr::select ( one_of ( c ( 'oln_id_a'))) %>%
	dplyr::rename( oln_id = oln_id_a) %>%
	distinct()

count(tf_list )

## count the number of substrates

targets_list <- tf_network_cleaned %>%
	dplyr::select ( one_of ( c ( 'oln_id_b'))) %>%
	dplyr::rename( oln_id = oln_id_b) %>%
	distinct()

targets_excluding_tf <- setdiff( targets_list, tf_list)

count( targets_excluding_tf)

## total number of proteins
length( unique( c ( targets_list$oln_id , tf_list$oln_id) ))

```

```{r}
# Odds ratio enrichment for transcriptional regulatory network.

fisher.test ( matrix (c(3156
,620
,26786
,2281), 2, 2))

1/0.4334889

```