### 2017-08-05 Source/Examples/rewire_network.Rmd

---
title: "rewire_network"
author: "Ignatius Pang"
date: "7 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(igraph)
library(tibble)
library(dplyr)
library(ggplot2)

 base_directory <- "/home/ignatius/PostDoc/2016/Triplet_Motifs/"
# base_directory <- "/home/ignatius/Documents/Postdoc/2016/Triplet_Motifs/"

data_directory <- file.path( base_directory, "Data/Triplet_Motifs_R_data")

load( file.path(data_directory, "network_data_library.Rdata"))

source( file.path(base_directory, "Source/Common", "count_triplet_motifs_helper.R"))

```

## R Markdown

```{r}

# Make a ring with 10 vertices 
g <- make_ring(10, circular = TRUE)

# Plot the ring 
plot(g)

# Count the number of edges in the graph
number_of_edges <- ecount( g)

# Rewire the graphs by swapping edges 10 times, keeping the degree distribution the same
 graph_1 <- g %>%
			rewire(keeping_degseq(loops = FALSE, niter = number_of_edges))

 				# loops : Whether to allow destroying and creating loop edges.
 				# niter : number of rewiring trials to perform

# Plot the graph
 plot(graph_1)

# What happens if I don't keep the edgree distribution the same 
 graph_2 <- g %>%
					 rewire( each_edge(prob=1, loops = FALSE, multiple=FALSE))

 				# loops : Whether to allow destroying and creating loop edges.
 				# multiple : Logical scalar, whether multiple edges are allowed int the generated graph

 plot(graph_2)

```

```{r }

format_degree_distribution_as_table <- function( igraph_object, label) {

	table_of_node_degrees <- table( degree(igraph_object))

	## Add 1 to the values to avoid log(0)
	x_axis <- 	as.numeric( names( table_of_node_degrees ) )+ 1
	y_axis <- as.numeric(t(table_of_node_degrees)) + 1

	# plot( x_axis , y_axis,  log="xy")

	head(y_axis)

	result_table <- data.frame( label=rep(label, length(x_axis)), degrees=x_axis, frequency=y_axis  )

	return(result_table)
}

## Obtain a protein-protein interactions network

head( sbi_interactome_collated)

# Read in the graph object
graph_1 <- graph.data.frame(sbi_interactome_collated[, c("oln_id_a", "oln_id_b")], directed=FALSE, vertices=NULL)

# Tabulate and graph the node degrees of the original network
graph_1_table <- format_degree_distribution_as_table(graph_1, "Original Network")

# Rewire the network using degree-preserving randomization  
# Plot the degree distribution of the network afterwards
graph_2 <- graph_1  %>%
			rewire(keeping_degseq(loops = FALSE, niter = number_of_edges))

graph_2_table <- format_degree_distribution_as_table(graph_2, "Rewired, keeping degree distribution")

# Rewire the network without keeping the degree distribution the same
# Plot the degree distribution of the network afterwards
graph_3 <- graph_1  %>%
			rewire(each_edge(prob=0.4, loops = FALSE, multiple=FALSE))

graph_3_table <- format_degree_distribution_as_table(graph_3, "Random rewiring")

merged_table <- rbind( graph_1_table, graph_2_table, graph_3_table)

ggplot( merged_table, aes( degrees, frequency)) +
	geom_point() +
	facet_grid ( . ~ label) + 
    scale_x_continuous(trans='log2') + 
    scale_y_continuous(trans='log2')

```