### 2018-01-03 Source/More_stringent_network/fitness_scores_quartiles.Rmd
Add files

---
title: "Analysis of the distribution of fitness scores for strigent set (negative interactions only)"
output: html_notebook
---

```{r}

plot (density( filtered_costanzo_stringent[, "genetic_interaction_score"]))

summary( filtered_costanzo_stringent[, "genetic_interaction_score"])

#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# -1.1616 -0.2308 -0.1670 -0.2077 -0.1381 -0.1201

filtered_costanzo_stringent  %>% dplyr::count()
dplyr::filter( filtered_costanzo_stringent,  genetic_interaction_score < -0.1381) %>% dplyr::count()
dplyr::filter( filtered_costanzo_stringent,  genetic_interaction_score < -0.1670) %>% dplyr::count()
dplyr::filter( filtered_costanzo_stringent,  genetic_interaction_score < -0.2308) %>% dplyr::count()

```

```{r}

quantile_threshold <- quantile( filtered_costanzo_stringent[, "genetic_interaction_score"], probs=seq(0.70, 0.95, 0.05))

names( quantile_threshold)

```