### 2018-01-19 Source/Supplementary_Files/percentage_of_network_analyzed.Rmd

---
title: "percentage_of_network_analyzed.Rmd"
author: "Ignatius Pang"
date: "31 August 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r}
library(dplyr)
library(tibble)
library(tidyr)
```

## Global Parameters
```{r}
local_base_directory <- "/home/ignatius/PostDoc/2016/Triplet_Motifs"
options <- commandArgs(trailingOnly = TRUE)

source( file.path( local_base_directory, "Source/Common/parameters_file.R") ) 
source( file.path ( source_directory_common, 'count_triplet_motifs_helper.R'))

# data_directory <- "/media/z3371724/PostDoc/2016/Triplet_Motifs/Data/Triplet_Motifs_R_data/"

load( paste(data_directory, "network_data_library.Rdata", sep=""))

triplet_motifs_costanzo <- readRDS( file=paste(data_directory, "triplet_motifs_costanzo_2016.Rdata", sep="") )

filtered_costanzo_stringent <- filtered_costanzo_stringent_2016

colnames( filtered_costanzo_stringent)

significant_types_of_motifs <- c("pp", "tutu" ,"pku" ,"kuku","tdp", "pkd"  )

```

## Get all unique negative genetic interactions

Fact check: The six types of triplet motifs are associated with 80% of the negative genetic interactions present in triplet motifs, and 8% of the known negative genetic interactions.

```{r}

all_negative_gi_helper_a <- filtered_costanzo_stringent[ , c("query_oln_id_edited", "array_oln_id_edited") ] %>%
							dplyr::rename( oln_id_a=query_oln_id_edited) %>%
							dplyr::rename( oln_id_b=array_oln_id_edited)

all_negative_gi_helper_b <- filtered_costanzo_stringent[ , c("query_oln_id_edited", "array_oln_id_edited") ] %>%
							dplyr::rename( oln_id_b=query_oln_id_edited) %>%
							dplyr::rename( oln_id_a=array_oln_id_edited)

all_negative_gi <- dplyr::union  (all_negative_gi_helper_a, all_negative_gi_helper_b ) %>%
					dplyr::distinct() %>%
				   dplyr::filter( oln_id_a >= oln_id_b ) 
```

## how many genes are in the negative genetic interaction network

```{r}
dplyr::union ( all_negative_gi[, "oln_id_b"], all_negative_gi[, "oln_id_a"] ) %>% 
	unique() %>% 
	length()

# as.data.frame(filtered_costanzo_stringent[, "query_oln_id_edited"])  %>% 
# 	rename( array_oln_id_edited=query_oln_id_edited) %>% 
# 	dplyr::union ( as.data.frame(filtered_costanzo_stringent[, "array_oln_id_edited"]) ) %>% 
# 	distinct() %>% 
# 	count()

```

## Get the unique genetic interactions in triplet motifs 
```{r}
all_gi_in_triplet_motifs_helper_a <- triplet_motifs_costanzo[, c("oln_id_a", "oln_id_b", "type_ac", "type_bc")]

all_gi_in_triplet_motifs_helper_b <- triplet_motifs_costanzo[, c("oln_id_b", "oln_id_a", "type_ac", "type_bc")] %>%
										dplyr::rename( temp=oln_id_b) %>%
										dplyr::rename( oln_id_b=oln_id_a)  %>%
										dplyr::rename( oln_id_a=temp)

all_gi_in_triplet_motifs <- dplyr::union ( all_gi_in_triplet_motifs_helper_a, 
										   all_gi_in_triplet_motifs_helper_b) %>%
							dplyr::filter( oln_id_a >= oln_id_b )  %>%
							dplyr::mutate( motif_type = ifelse( type_ac >= type_bc,
																paste (type_ac, type_bc, sep="" ) ,  
																paste (type_bc, type_ac, sep="" ) ) ) %>%
							dplyr::select(one_of( c("oln_id_a", "oln_id_b", "motif_type")))

## In significant triplet motifs 
all_gi_in_sig_triplet_motifs <- all_gi_in_triplet_motifs %>%
								dplyr::filter ( motif_type %in% significant_types_of_motifs ) %>%
								dplyr::select(one_of( c("oln_id_a", "oln_id_b"))) %>%
								dplyr::distinct()

## In all triplet motifs
all_gi_in_triplet_motifs <- all_gi_in_triplet_motifs %>%
							dplyr::select(one_of( c("oln_id_a", "oln_id_b"))) %>%
							dplyr::distinct()

```

## Perform counting 
```{r}
all_gi_in_sig_triplet_motifs_count_check <- dplyr::count( dplyr::intersect ( all_gi_in_sig_triplet_motifs, all_negative_gi))[[1]]

all_gi_in_sig_triplet_motifs_count <- dplyr::count ( all_gi_in_sig_triplet_motifs)[[1]]

if( all_gi_in_sig_triplet_motifs_count_check != all_gi_in_sig_triplet_motifs_count) {
	stop ('Something is wrong with the counts for gi in triplet motifs.')
}

all_gi_in_triplet_motifs_counts <- dplyr::count(all_gi_in_triplet_motifs)[[1]]

all_negative_gi_counts <- dplyr::count( all_negative_gi)[[1]]

print ( "Number of unique negative genetic interactions in total: ")
all_negative_gi_counts

print ( "Number of unique negative genetic interactions in all 15 triplet motifs: ")
all_gi_in_triplet_motifs_counts

print ( "Number of unique genetic interactions in the six overrepresented triplet motifs: ")
all_gi_in_sig_triplet_motifs_count

```

## Calculate fractions
```{r}

print ( "Proportion of unique negative genetic interactions in all 15 triplet motifs: ")
all_gi_in_triplet_motifs_counts/all_negative_gi_counts*100

print ( "Proportion of unique negative genetic interactions in the six overrepresented triplet motifs: ")
all_gi_in_sig_triplet_motifs_count/all_negative_gi_counts*100

print( "Proportion of the unique negative genetic interactions in triplet motifs that are part of significant triplet motifs: ")
all_gi_in_sig_triplet_motifs_count/all_gi_in_triplet_motifs_counts*100
```

## There was one triplet motif that was of the type TUKU
```{r}

triplet_motifs_costanzo %>% 
	dplyr::filter( ( type_ac == 'tu' & type_bc == 'ku' ) |
				   	( type_ac == 'ku' & type_bc == 'tu' )  )

```

## Count negative genetic interactions where either gene is essential
```{r}

essential_genes_edited <- essential_genes %>% dplyr::mutate ( is_essential=1)

## No. of all negative genetic interactions in which either one of the interacting genes are essential
count_ngi_essential <- all_negative_gi %>%
	dplyr::left_join ( essential_genes_edited, by=c("oln_id_a" = "oln_id")) %>%
	dplyr::left_join ( essential_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_essential.x == 1 | is_essential.y == 1) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_essential

## No. of all negative genetic interactions in triplet motifs in which either one of the interacting genes are essential
count_ngi_triplets_essential <- all_gi_in_triplet_motifs %>%
	dplyr::left_join ( essential_genes_edited, by=c("oln_id_a" = "oln_id")) %>%
	dplyr::left_join ( essential_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_essential.x == 1 | is_essential.y == 1) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_triplets_essential

## No. of all negative genetic interactions in significant triplet motifs in which either one of the interacting genes are essential
count_ngi_sig_triplets_essential <- all_gi_in_sig_triplet_motifs  %>%
	dplyr::left_join ( essential_genes_edited, by=c("oln_id_a" = "oln_id")) %>%
	dplyr::left_join ( essential_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_essential.x == 1 | is_essential.y == 1) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_sig_triplets_essential

## No. of all negative genetic interactions in triplet motifs but neither of the protein in the interacting pair are essential
count_ngi_triplets_not_essential <- all_gi_in_triplet_motifs %>%
	dplyr::setdiff ( all_gi_in_triplet_motifs %>%
	dplyr::left_join ( essential_genes_edited, by=c("oln_id_a" = "oln_id")) %>%
	dplyr::left_join ( essential_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_essential.x == 1 | is_essential.y == 1) %>%
	dplyr::select ( oln_id_a, oln_id_b)) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_triplets_not_essential

## Fishers exact test
# Does the negative interacting pair contain at least one protein that is essential = TRUE, is the negative genetic interaction within triplet = TRUE
essential_fisher_exact_test_a <- count_ngi_triplets_essential

# Does the negative interacting pair contain at least one protein that is essential = FALSE, is the negative genetic interaction within triplet = TRUE
essential_fisher_exact_test_b <- count_ngi_triplets_not_essential

# Does the negative interacting pair contain at least one protein that is essential = TRUE, is the negative genetic interaction within triplet = FALSE
essential_fisher_exact_test_c <- count_ngi_essential - count_ngi_triplets_essential

# Does the negative interacting pair contain at least one protein that is essential = FALSE, is the negative genetic interaction within triplet = FALSE

all_genetic_interactions_count <- (all_negative_gi %>% dplyr::count() %>% as.numeric())
essential_fisher_exact_test_d <- all_genetic_interactions_count - essential_fisher_exact_test_a - essential_fisher_exact_test_b - essential_fisher_exact_test_c

essential_matrix <- matrix( c( essential_fisher_exact_test_a, essential_fisher_exact_test_b, 
						       essential_fisher_exact_test_c, essential_fisher_exact_test_d), 2, 2 )

essential_matrix

fisher.test ( essential_matrix)

```

## Count negative genetic interactions where either gene is toxic upon overexpression

```{r}

overexpressed_toxic_genes_edited <- overexpressed_toxic_genes %>% dplyr::mutate ( is_toxic_upon_ox =1)

## No. of all negative genetic interactions in which either one of the interacting genes are toxic upon overexpression
count_ngi_toxic <- all_negative_gi %>%
	dplyr::left_join ( overexpressed_toxic_genes_edited, by=c("oln_id_a" = "oln_id"))%>%
	dplyr::left_join ( overexpressed_toxic_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_toxic_upon_ox.x == 1 | is_toxic_upon_ox.y == 1) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_toxic

## No. of all negative genetic interactions in triplet motifs in which either one of the interacting genes are toxic upon overexpression
count_ngi_triplets_toxic <- all_gi_in_triplet_motifs %>%
	dplyr::left_join ( overexpressed_toxic_genes_edited, by=c("oln_id_a" = "oln_id"))%>%
	dplyr::left_join ( overexpressed_toxic_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_toxic_upon_ox.x == 1 | is_toxic_upon_ox.y == 1) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_triplets_toxic

## No. of all negative genetic interactions in significant triplet motifs in which either one of the interacting genes are toxic upon overexpression
count_ngi_sig_triplets_toxic <- all_gi_in_sig_triplet_motifs  %>%
	dplyr::left_join ( overexpressed_toxic_genes_edited, by=c("oln_id_a" = "oln_id"))%>%
	dplyr::left_join ( overexpressed_toxic_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_toxic_upon_ox.x == 1 | is_toxic_upon_ox.y == 1) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_sig_triplets_toxic

## No. of all negative genetic interactions in triplet motifs but neither of the proteins in the interacting pair are toxic upon overexpression
count_ngi_triplets_not_toxic <- all_gi_in_triplet_motifs %>%
	dplyr::setdiff ( all_gi_in_triplet_motifs %>%
	dplyr::left_join ( overexpressed_toxic_genes_edited, by=c("oln_id_a" = "oln_id"))%>%
	dplyr::left_join ( overexpressed_toxic_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_toxic_upon_ox.x == 1 | is_toxic_upon_ox.y == 1) %>%
	dplyr::select ( oln_id_a, oln_id_b)) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_triplets_not_toxic

## Fishers exact test
# Does the negative interacting pair contain at least one protein that is toxic upon overexpression = TRUE, is the negative genetic interaction within triplet = TRUE
toxic_fisher_exact_test_a <- count_ngi_triplets_toxic

# Does the negative interacting pair contain at least one protein that is toxic upon overexpression = FALSE, is the negative genetic interaction within triplet = TRUE
toxic_fisher_exact_test_b <- count_ngi_triplets_not_toxic

# Does the negative interacting pair contain at least one protein that is toxic upon overexpression = TRUE, is the negative genetic interaction within triplet = FALSE
toxic_fisher_exact_test_c <- count_ngi_toxic - count_ngi_triplets_toxic

# Does the negative interacting pair contain at least one protein that is toxic upon overexpression = FALSE, is the negative genetic interaction within triplet = FALSE

all_genetic_interactions_count <- (all_negative_gi %>% dplyr::count() %>% as.numeric())
toxic_fisher_exact_test_d <- all_genetic_interactions_count - toxic_fisher_exact_test_a - toxic_fisher_exact_test_b - toxic_fisher_exact_test_c

toxic_matrix <- matrix( c( toxic_fisher_exact_test_a, toxic_fisher_exact_test_b, 
						   toxic_fisher_exact_test_c, toxic_fisher_exact_test_d), 2, 2 )

toxic_matrix

fisher.test ( toxic_matrix)

```

## Count negative genetic interactions where either gene is cell-cycle regulated
```{r}

periodically_expressed_genes_edited <- periodically_expressed_genes %>% dplyr::mutate ( is_periodic =1)

## No. of all negative genetic interactions in which either one of the interacting genes are cell-cycle regulated
count_ngi_cell_cycle <- all_negative_gi %>%
	dplyr::left_join ( periodically_expressed_genes_edited, by=c("oln_id_a" = "oln_id"))%>%
	dplyr::left_join ( periodically_expressed_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_periodic.x == 1 | is_periodic.y == 1) %>%
	dplyr::count()  %>%
	as.numeric()

count_ngi_cell_cycle

## No. of all negative genetic interactions in triplet motifs in which either one of the interacting genes are cell-cycle regulated
count_ngi_triplets_cell_cycle <- all_gi_in_triplet_motifs %>%
	dplyr::left_join ( periodically_expressed_genes_edited, by=c("oln_id_a" = "oln_id"))%>%
	dplyr::left_join ( periodically_expressed_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_periodic.x == 1 | is_periodic.y == 1) %>%
	dplyr::count()  %>%
	as.numeric()

count_ngi_triplets_cell_cycle

## No. of all negative genetic interactions in significant triplet motifs in which either one of the interacting genes are cell-cycle regulated
count_ngi_sig_triplets_cell_cycle <- all_gi_in_sig_triplet_motifs  %>%
	dplyr::left_join ( periodically_expressed_genes_edited, by=c("oln_id_a" = "oln_id"))%>%
	dplyr::left_join ( periodically_expressed_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_periodic.x == 1 | is_periodic.y == 1) %>%
	dplyr::count()  %>%
	as.numeric()

count_ngi_sig_triplets_cell_cycle

## No. of all negative genetic interactions in triplet motifs but neither of the proteins in interacting pair are cell-cycle regulated
count_ngi_triplets_not_cell_cycle <- all_gi_in_triplet_motifs %>%
	dplyr::setdiff ( all_gi_in_triplet_motifs %>%
	dplyr::left_join ( periodically_expressed_genes_edited, by=c("oln_id_a" = "oln_id"))%>%
	dplyr::left_join ( periodically_expressed_genes_edited, by=c("oln_id_b" = "oln_id")) %>%
	dplyr::filter ( is_periodic.x == 1 | is_periodic.y == 1) %>%
	dplyr::select ( oln_id_a, oln_id_b)) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_triplets_not_cell_cycle

## Fishers exact test

# Does the negative interacting pair contain at least one periodically expressed gene = TRUE, is the negative genetic interaction within triplet = TRUE
cell_cycle_fisher_exact_test_a <- count_ngi_triplets_cell_cycle

# Does the negative interacting pair contain at least one periodically expressed gene = FALSE, is the negative genetic interaction within triplet = TRUE
cell_cycle_fisher_exact_test_b <- count_ngi_triplets_not_cell_cycle

# Does the negative interacting pair contain at least one periodically expressed gene = TRUE, is the negative genetic interaction within triplet = FALSE
cell_cycle_fisher_exact_test_c <- count_ngi_cell_cycle - count_ngi_triplets_cell_cycle

# Does the negative interacting pair contain at least one periodically expressed gene = FALSE, is the negative genetic interaction within triplet = FALSE
all_genetic_interactions_count <- (all_negative_gi %>% dplyr::count() %>% as.numeric())

cell_cycle_fisher_exact_test_d <- all_genetic_interactions_count - cell_cycle_fisher_exact_test_a - cell_cycle_fisher_exact_test_b - cell_cycle_fisher_exact_test_c

cell_cycle_matrix <- matrix( c( cell_cycle_fisher_exact_test_a, cell_cycle_fisher_exact_test_b, 
						        cell_cycle_fisher_exact_test_c, cell_cycle_fisher_exact_test_d), 2, 2 )

cell_cycle_matrix

fisher.test ( cell_cycle_matrix)

```

## Count negative genetic interactions where both genes are paralogs or ohnologs

```{r}

orthomcl_paralogs <- clean_graph_table  (orthomcl_paralogs,  "oln_id_a", "oln_id_b", directed=FALSE )

sgd_paralogs 	  <- clean_graph_table  (sgd_paralogs,  "oln_id_a", "oln_id_b", directed=FALSE )

orthomcl_paralogs <- dplyr::distinct ( dplyr::union ( orthomcl_paralogs, sgd_paralogs) )

orthomcl_paralogs <- clean_graph_table  (orthomcl_paralogs,  "oln_id_a", "oln_id_b", directed=FALSE )

orthomcl_paralogs_edited <- orthomcl_paralogs %>% dplyr::mutate ( is_paralog =1)

num_of_paralogs_pair <- orthomcl_paralogs_edited %>% dplyr::distinct() %>% dplyr::count() %>% as.numeric()

num_of_paralogs_pair

## No. of all negative genetic interactions in which the interacting pair are paralogs or ohnologs
count_ngi_paralogs <- all_negative_gi %>%
	dplyr::left_join ( orthomcl_paralogs_edited , by = c( "oln_id_a" = "oln_id_b", "oln_id_b" = "oln_id_a")) %>%
	dplyr::left_join ( orthomcl_paralogs_edited , by = c( "oln_id_b" = "oln_id_b", "oln_id_a" = "oln_id_a")) %>%
	dplyr::filter ( is_paralog.x == 1 | is_paralog.y==1 ) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_paralogs

## No. of all negative genetic interactions in triplet motifs in which the interacting pair are paralogs or ohnologs
count_ngi_triplets_paralogs <- all_gi_in_triplet_motifs %>%
	dplyr::left_join ( orthomcl_paralogs_edited , by = c( "oln_id_a" = "oln_id_b", "oln_id_b" = "oln_id_a"))%>%
	dplyr::left_join ( orthomcl_paralogs_edited , by = c( "oln_id_b" = "oln_id_b", "oln_id_a" = "oln_id_a"))%>%
	dplyr::filter ( is_paralog.x == 1 | is_paralog.y==1 ) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_triplets_paralogs

## No. of all negative genetic interactions in the six significant triplet motifs in which the interacting pair are paralogs or ohnologs
count_ngi_sig_triplets_paralogs <- all_gi_in_sig_triplet_motifs  %>%
	dplyr::left_join ( orthomcl_paralogs_edited , by = c( "oln_id_a" = "oln_id_b", "oln_id_b" = "oln_id_a"))%>%
	dplyr::left_join ( orthomcl_paralogs_edited , by = c( "oln_id_b" = "oln_id_b", "oln_id_a" = "oln_id_a"))%>%
	dplyr::filter ( is_paralog.x == 1 | is_paralog.y==1 ) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_sig_triplets_paralogs

## No. of all negative genetic interactions in triplet motifs but the interacting pair are not paralogs or ohnologs
count_ngi_triplets_not_paralogs <- all_gi_in_triplet_motifs %>%
	dplyr::setdiff ( all_gi_in_triplet_motifs %>%
	dplyr::left_join ( orthomcl_paralogs_edited , by = c( "oln_id_a" = "oln_id_b", "oln_id_b" = "oln_id_a"))%>%
	dplyr::left_join ( orthomcl_paralogs_edited , by = c( "oln_id_b" = "oln_id_b", "oln_id_a" = "oln_id_a"))%>%
	dplyr::filter ( is_paralog.x == 1 | is_paralog.y ==1 ) %>%
	dplyr::select ( oln_id_a, oln_id_b)) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_triplets_not_paralogs

## Fishers exact test

# Does the negative interacting pair contain a paralog = TRUE, is the negative genetic interaction within triplet = TRUE
paralog_fisher_exact_test_a <- count_ngi_triplets_paralogs

# Does the negative interacting pair contain a paralog = FALSE, is the negative genetic interaction within triplet = TRUE
paralog_fisher_exact_test_b <- count_ngi_triplets_not_paralogs

# Does the negative interacting pair contain a paralog = TRUE, is the negative genetic interaction within triplet = FALSE
paralog_fisher_exact_test_c <- count_ngi_paralogs - count_ngi_triplets_paralogs

# Does the negative interacting pair contain a paralog = FALSE, is the negative genetic interaction within triplet = FALSE
all_genetic_interactions_count <- (all_negative_gi %>% dplyr::count() %>% as.numeric())
paralog_fisher_exact_test_d <- all_genetic_interactions_count - paralog_fisher_exact_test_a - paralog_fisher_exact_test_b - paralog_fisher_exact_test_c

paralog_matrix <- matrix( c( paralog_fisher_exact_test_a, paralog_fisher_exact_test_b, 
						  paralog_fisher_exact_test_c, paralog_fisher_exact_test_d), 2, 2 )

paralog_matrix

fisher.test ( paralog_matrix)

```

## Negative genetic interactions overlapping with other types of interactions
```{r}

all_interactions <- read.table ( "/home/ignatius/PostDoc/2016/Triplet_Motifs/Results/Supplementary_Files/all_interactions.tab", 
 								 header=TRUE, sep="\t", quote="", stringsAsFactors = FALSE)

# Total number of negative genetic interactions
all_interactions %>%
	dplyr::filter( oln_id_a > oln_id_b) %>%
	dplyr::select(one_of( c( "oln_id_a", "oln_id_b", "interaction_type_abbrev" ))) %>%
	dplyr::mutate( count = 1) %>%
	tidyr::spread( "interaction_type_abbrev", "count") %>%
	dplyr::filter( ifelse ( is.na(ng), FALSE,  ng == 1) ) %>%
	dplyr::count() %>%
	as.numeric()

all_genetic_interactions_count <- (all_negative_gi %>% dplyr::count() %>% as.numeric())

# Total number of negative genetic interactions which overlap with at least one other type of interaction 
count_ngi_overlap <- all_interactions %>%
	dplyr::filter( oln_id_a > oln_id_b) %>%
	dplyr::select(one_of( c( "oln_id_a", "oln_id_b", "interaction_type_abbrev" ))) %>%
	dplyr::mutate( count = 1) %>%
	tidyr::spread( "interaction_type_abbrev", "count") %>%
	dplyr::filter(   ifelse ( is.na(ng), FALSE,  ng == 1) & 
					 	( ifelse ( is.na(kd), FALSE, kd >= 1) |
					 	  ifelse ( is.na(ku), FALSE, ku >= 1) |
					 	  ifelse ( is.na(p ), FALSE, p  >= 1) |
					 	  ifelse ( is.na(td), FALSE, td >= 1) |
					 	  ifelse ( is.na(tu), FALSE, tu >= 1)  ) ) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_overlap

# Total number of interactions, including all types of interactions. Overlapping interactions were only counted once.
count_ngi_triplets_overlap <- all_interactions %>%
	dplyr::filter( oln_id_a > oln_id_b & is_part_of_triplet_motif == 1) %>%
	dplyr::select(one_of( c( "oln_id_a", "oln_id_b", "interaction_type_abbrev" ))) %>%
	dplyr::mutate( count = 1) %>%
	tidyr::spread( "interaction_type_abbrev", "count") %>%
	dplyr::filter( ng == 1 & ( 	kd == 1 | ku == 1 | p == 1 | td == 1 | tu == 1) ) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_triplets_overlap

# Total number of interactions where , including all types of interactions. Overlapping interactions were only counted once.
count_ngi_sig_triplets_overlap <- all_interactions %>%
	dplyr::inner_join(all_gi_in_sig_triplet_motifs ) %>%
	dplyr::filter( oln_id_a > oln_id_b & is_part_of_triplet_motif == 1) %>%
	dplyr::select(one_of( c( "oln_id_a", "oln_id_b", "interaction_type_abbrev" ))) %>%
	dplyr::mutate( count = 1) %>%
	tidyr::spread( "interaction_type_abbrev", "count") %>%
	dplyr::filter( ng == 1 & ( 	kd == 1 | ku == 1 | p == 1 | td == 1 | tu == 1) ) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_sig_triplets_overlap

## No. of all negative genetic interactions in triplet motifs but the interacting pair do not overlap with other types of interactions 
count_ngi_triplets_not_overlap <- all_gi_in_triplet_motifs %>%
	dplyr::setdiff ( all_interactions %>%
	dplyr::filter( oln_id_a > oln_id_b & is_part_of_triplet_motif == 1) %>%
	dplyr::select(one_of( c( "oln_id_a", "oln_id_b", "interaction_type_abbrev" ))) %>%
	dplyr::mutate( count = 1) %>%
	tidyr::spread( "interaction_type_abbrev", "count") %>%
	dplyr::filter( ng == 1 & ( 	kd == 1 | ku == 1 | p == 1 | td == 1 | tu == 1) ) %>%
	dplyr::select ( oln_id_a, oln_id_b)) %>%
	dplyr::count() %>%
	as.numeric()

count_ngi_triplets_not_overlap

## Fishers exact test

# Does the negative interacting pair ovelap with other types of interactions = TRUE, is the negative genetic interaction within triplet = TRUE
overlap_fisher_exact_test_a <- count_ngi_triplets_overlap

# Does the negative interacting pair ovelap with other types of interactions = FALSE, is the negative genetic interaction within triplet = TRUE
overlap_fisher_exact_test_b <- count_ngi_triplets_not_overlap

# Does the negative interacting pair ovelap with other types of interactions = TRUE, is the negative genetic interaction within triplet = FALSE
overlap_fisher_exact_test_c <- count_ngi_overlap - count_ngi_triplets_overlap

# Does the negative interacting pair ovelap with other types of interactions = FALSE, is the negative genetic interaction within triplet = FALSE
all_genetic_interactions_count <- (all_negative_gi %>% dplyr::count() %>% as.numeric())
overlap_fisher_exact_test_d <- all_genetic_interactions_count - overlap_fisher_exact_test_a - overlap_fisher_exact_test_b - overlap_fisher_exact_test_c

overlap_matrix <- matrix( c( overlap_fisher_exact_test_a, overlap_fisher_exact_test_b, 
						  overlap_fisher_exact_test_c, overlap_fisher_exact_test_d), 2, 2 )

overlap_matrix

fisher.test ( overlap_matrix)

```

## Calculating the number of paralog pairs that are int tiplets and show a negative interaction: 
I'll need to calculate the number of interactions that overlap with pairs of paralogous proteins. 
```{r}
triplet_motifs_costanzo_2016 <- read.table ( file.path( data_directory, "triplet_motifs_costanzo_2016.txt"), 
											 header =TRUE , stringsAsFactors = FALSE)

gene_a <- as.character( as.vector(c( t(triplet_motifs_costanzo_2016[,1]) ,
									 t(triplet_motifs_costanzo_2016[,1]) , 
									 t(triplet_motifs_costanzo_2016[,2]) ) ))

gene_b <- as.character( as.vector(c( t(triplet_motifs_costanzo_2016[,2]) ,
									 t(triplet_motifs_costanzo_2016[,3])    , 
									 t(triplet_motifs_costanzo_2016[,3]) ) ))

edge_type_attributes <- as.character( as.vector( c( 
		rep( "ng", length(triplet_motifs_costanzo_2016[,1] )),
		t(triplet_motifs_costanzo_2016[, c("type_ac")]),
		t(triplet_motifs_costanzo_2016[, c("type_bc")])
	)  ) )

interactions_from_triplet_motifs_left <- data.frame( oln_id_a = gene_a, oln_id_b = gene_b, type_of_interaction = edge_type_attributes  )

interactions_from_triplet_motifs_right <- data.frame( oln_id_a = gene_b, oln_id_b = gene_a, type_of_interaction = edge_type_attributes  )

interactions_from_triplet_motifs <- interactions_from_triplet_motifs_left %>%
	dplyr::union( interactions_from_triplet_motifs_right) %>%
	dplyr::filter ( oln_id_a > oln_id_b ) %>%
	dplyr::distinct()

interactions_match_paralogs <-  interactions_from_triplet_motifs %>%
								dplyr::left_join( orthomcl_paralogs_edited, by = c( "oln_id_a" = "oln_id_b", "oln_id_b" = "oln_id_a")) %>%
								dplyr::left_join( orthomcl_paralogs_edited, by = c( "oln_id_b" = "oln_id_b", "oln_id_a" = "oln_id_a")) %>%
								dplyr::filter( is_paralog.x ==1 | is_paralog.y==1 ) %>%
								dplyr::distinct()

# all unique pairs of paralogs overlapping biological interactions
paralogs_unique_pairs_overlap_interactions <- interactions_match_paralogs %>%
												dplyr::select ( one_of( c( "oln_id_a", "oln_id_b"))) %>%
												dplyr::distinct()

dplyr::count(paralogs_unique_pairs_overlap_interactions) # 101 pairs

## What do the ones that are not negative genetic interactions consists of:
negative_genetic_interaction_paralog <- interactions_match_paralogs %>% 
										dplyr::filter ( type_of_interaction != "ng" ) %>% 	 
										dplyr::select ( one_of( c( "oln_id_a", "oln_id_b"))) %>%
										dplyr::distinct()

dplyr::count(negative_genetic_interaction_paralog) # 38 rows

other_biological_interactions_paralog <- interactions_match_paralogs %>% 
										 dplyr::filter ( type_of_interaction == "ng" )  %>% 	
										 dplyr::select ( one_of( c( "oln_id_a", "oln_id_b"))) %>%
										 dplyr::distinct()

dplyr::count( other_biological_interactions_paralog ) # 80 rows

# Paris of Paralogs that form negative genetic interactions and other biological interactions: 17 rows
dplyr::intersect( negative_genetic_interaction_paralog, other_biological_interactions_paralog)

80 + 38 - 17 == 101

```